### 2023-12-18 draft_jss/draft_jss.Rmd
update draft

---
documentclass: jss
bibliography: article.bib
author:
  - name: FirstName LastName
    orcid: 0000-0000-0000-0000
    affiliation: University/Company
    # use this syntax to add text on several lines
    address: |
      | First line
      | Second line
    email: \email{name@company.com}
    url: https://posit.co
  - name: Second Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation \AND'
    # To add another line, use \AND at the end of the previous one as above
  - name: Third Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation'
    # To add another line, use \AND at the end of the previous one as above
  - name: Third Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation \AND'
    # To add another line, use \AND at the end of the previous one as above
  - name: Third Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation'
    # To add another line, use \AND at the end of the previous one as above
  - name: Fourth Author
    orcid: 0000-0000-0000-0000
    address: |
      | Department of Statistics and Mathematics,
      | Faculty of Biosciences,
      | Universitat Autònoma de Barcelona
    affiliation: |
      | Universitat Autònoma 
      | de Barcelona
    # use a different affiliation in adress field (differently formated here)
    affiliation2: Universitat Autònoma de Barcelona
title:
  formatted: "Reservoir computing in R : a tutorial using \\pkg{reservoirnet}"
  # If you use tex in the formatted title, also supply version without
  plain:     "Reservoir computing in R : a tutorial using reservoirnet"
  # For running headers, if needed
  short:     "\\pkg{reservoirnet}: reservoir computing in R"
abstract: >
  The abstract of the article.
keywords:
  # at least one keyword must be supplied
  formatted: [keywords, not capitalized, "\\proglang{Java}"]
  plain:     [keywords, not capitalized, Java]
preamble: >
  \usepackage{amsmath}
output:
  bookdown::pdf_book:
    base_format: rticles::jss_article
  bookdown::word_document2:
    default
knit: (function(inputFile, encoding){
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = here::here("draft_jss"), output_format = "all") })
---

```{r, setup, include=FALSE}
options(prompt = 'R> ', continue = '+ ')
set.seed(1)
```

# Introduction

Reservoir Computing (RC) is a prominent machine learning method, proposed by @jaeger__2001, @maass_real-time_2002 and @lukosevicius_reservoir_2009
that has gained significant attention in recent years for its ability to
efficiently process information generated by dynamical systems. This
innovative approach leverages the dynamics of a high-dimensional
"reservoir" (we define it below) to perform complex computations and solve various tasks based
on the response of this dynamical system to input signals. RC has demonstrated its efficacy in tackling various challenges, encompassing pattern classification and time series forecasting in applications ranging from electrocardiogram analysis to bird calls @trouvain_canary_2021, language processing @hinaut_real-time_2013, power plants, internet traffic, stock prices, and beyond [@tanaka_recent_2019, @lukosevicius_reservoir_2009].

Originally, the RC paradigm was implemented in artificial firing-rate neurons  (“Echo State Networks”, @jaeger__2001) and spiking neurons ("Liquid State Machine”, @maass_real-time_2002) as a recurrent neural network (RNN) where the internal recurrent connections, denoted as the reservoir, are randomly generated and only the output layer (named “read-out”) is trained. The reservoir projects temporal input signals onto a high-dimensional feature space facilitating the learning of non-linear and temporal interactions. Thus, this recurrent layer contains are high-dimensional non-linear recombination of the inputs and past states: it is a “reservoir of computations” from which useful information can be linearly extracted (or “read-out”) to provide the desired outputs. This offers the advantage of decreasing the computing time while consistently maintaining robust performance compared to conventional RNNs [@vlachas_backpropagation_2020]. In parallel, this RC paradigm gained an increasing interest because of its ability to be implemented not only on classical computers. Indeed, the hidden recurrent layer can be kept untrained and thus a wide range of physical media can be used to replace it. Recently, @tanaka_recent_2019 reviewed this prolific field: from FPGA hardware [@penkovsky_efficient_2018], to spin waves using magnetic properties [@nakane_reservoir_2018], skrymions [@prychynenko_magnetic_2018] or optical implementations [@rafayelyan_large-scale_2020]. This provides interesting and potentially more efficient alternative to traditional machine learning.

RC leverages various hyperparameters to introduce prior knowledge about the relationship between input variables and output targets. However, since the connections within the reservoir are randomly initialized, the same set of hyperparameters may exhibit diverse behaviors across different realizations of the reservoir connections. This unpredictability makes it challenging to anticipate the performance of a particular hyperparameter setting, as identical settings may produce varying outcomes when applied to distinct instances of the reservoir. Moreover, selecting the most suitable hyperparameters often requires researchers to experiment with multiple combinations on a training dataset and evaluate their performance on a separate test set. Although this approach can be resource-intensive and time-consuming, it is a compromise that is acceptable considering the rapid simulation capabilities offered by RC.

*XH : Commentaires généraux sur le paragraphe : il faut ne faut pas insister sur les hyperparamètres comme si c’était négatif. Plutôt dire qu’il y a des optimisations et tricks à connaitre, et que c’est intéressant d’avoir une librairie pour ça, et aussi à des fins de reproductibilité. Dire que c’est un trade-off car c’est les a priori nécessaire qu’on rajoute dans l’algo car on gagne du temps sur l’apprentissage. C’est ambigu de parler de stochaticité ici je trouve, car le processus n’est pas stochastic une fois les poids fixés.*

*TF : j'ai modifié, mieux ?*

The utilization of RC poses challenges on multiple fronts. Firstly, its stochastic nature, attributed to the random aspect of the reservoir, introduces unpredictability into the results, even with well-chosen hyperparameters. Secondly, RC relies on numerous hyperparameters to establish connections between neurons. Guidance on determining these hyperparameters is limited, often requiring users to evaluate the performance of different combinations of hyperparameters on a train set before selecting the optimal combination to forecast on the test set^[In this article, we employ the term "train set" to refer to the combined dataset consisting of both the training and validation sets, which are cycled through in a cross-validation manner.]. Thirdly, there is a current absence of implementation in R, rendering the method challenging for users unfamiliar with Python [@trouvain_reservoirpy_2022] or Julia [@martinuzzi_reservoircomputingjl_2022].

*XH : Je pense que ça embrouille là : c’est train/test/validation Les hyperparamètres sont entrainé sur le train set puis évalués sur le validation set. Et ensuite le test set. Par abus de langage on dit souvent « test set » au lieu de « validation set » quand on optimise les paramètres. Moi ça me choque pas, mais dans un papier vaut mieux utiliser les termes adéquats, ou en tous cas toujours utiliser les memes, donc si tu n’as jamais de validation set (enfin le vrai test set), ça peut passer (pas besoin de tout changer). Met juste une note de bas de page pour préciser que en général on devrait l’appeler le validation set. On verra ce que les reviewers disent, pas besoin de perdre du temps maintenant.*

*TF : ok, j'ai rajouté une note de bas de page ci-dessous*

In this paper, we endeavor to delve into these challenges, offering comprehensive guidance to assist new users in maximizing the benefits of RC. Initially, a broad introduction to reservoir computing is presented, accompanied by a tutorial on its application using `reservoirnet`, an R package built upon the `reservoirPy` Python module [@trouvain_reservoirpy_2020, @trouvain_create_2022, @trouvain_reservoirpy_2022]. Subsequently, we investigate the various challenges associated with the use of RC for forecasting COVID-19 hospitalizations. This exploration includes detailed guidance on the modeling strategy, the selection of hyperparameters, and the implementation process.

# RC presentation

RC is a machine learning paradigm which is most often implemented as Echo State Networks (ESNs; @jaeger__2001), the firing-rate neuron version. An ESN is described by three matrices of connectivity: an input layer $W_{in}$, a recurrent layer $W$ and an output layer $W_{out}$. At each time step, the input vector $u_t$ is projected into the reservoir which is also combined with reservoir past state $x(t-1)$ through the recurrent connections. The output $y(t)$ is linearly read-out from the reservoir. Input $W_{in}$ and recurrent $W$ matrices are kept random; only the output matrix $W_{out}$ is trained in an offline or online method. Often a ridge regression (i.e. a regularized linear regression) is used to obtain the desired outputs $y(t)$ from the reservoir states $x(t)$. Figure \@ref(fig:rcpresentation) depicts the architecture. For simplicity, we will use the term "reservoir computing” for “Echo State Network” in the remainder of the paper.

```{r rcpresentation, echo=FALSE, fig.cap="my caption", out.width = '90%', fig.cap="Reservoir computing is composed of an input layer, a reservoir and an output layer. Connection between input layer and reservoir and inside reservoir are random. Only the output layer is optimized based on a ridge regression."}
knitr::include_graphics("images/schema.png/image1.png")
```

The input layer $u(t)$ is an $M$-dimension vector, where $M$ is the number of input time series, which corresponds to the values of the input time series at time $t$ where $t = 1, …, T$. The reservoir layer $x(t)$ is an $N_{res}$-dimensional vector where $N_{res}$ is the number of nodes in the reservoir. The value $x(t)$ is defined as follow:

$$x( t+1 ) = ( 1 - \alpha )  x ( t) + \alpha \: tanh( W x(t) + W_{in} u(t+1) ) \text{ , where } \alpha \in [0, 1 ]$$

The leaking rate $\alpha$ defines the rate of update of the nodes. The
closer $\alpha$ is to $1$, the higher the reservoir is sensitive to new
inputs (i.e $u(t)$). Therefore, the reservoir state at time $t+1$
denoted $x(t+1)$ depends on the reservoir state at the previous time
(i.e $x(t)$) and the new inputs (i.e $u(t+1)$). Both $W_{in}$ and $W$
are random matrices of size $N_{res} \times M$ and $N_{res} \times N_{res}$
respectively.

The leaking rate $\alpha$ defines the update rate of the nodes. The closer $\alpha$ is to $1$, the more the reservoir is sensitive to new inputs $u(t)$. Therefore, the reservoir state at time $t+1$ denoted $x(t+1)$ depends on the reservoir state at the previous time $x(t)$ and the new inputs $u(t+1)$. Both $W_{in}$ and $W$ are random matrices of size $N_{res} \times M$ and $N_{res} \times N_{res}$ respectively.

*XH : Ah tu utilises une matrice dense ? si oui, t’as fait ça dans toutes tes simulations ? habituellement on utilise plutôt 10 ou 20% de connectivité, surtout si on prend un distribution bimodale comme bernouilli. En tous cas il faut pas préciser que soit doit être dense ici.*

*TF : on a pris les paramètres par défaut de reservoirpy, si j'ai bien compris, la matrice est générée par la fonction reservoirpy.mat_gen.bernoulli qui par défaut est dense (connectivity = 1)*

$W_{in}$ is a dense matrix generated using a Bernoulli (bimodal) distribution where each value can be either $-I_{scale}(m)$ or $I_{scale}(m)$ with an equal probability where  $m = 1, …, M$ corresponds to a given feature in the input layer. The input scaling, denoted $I_{scale}$, is a hyperparameter coefficient often common to all features from the input layer or specific to each feature $m$. In that case, the more important the feature is, the greater should be its input scaling. $W$ is a matrix (usually sparse) where values are generated from a Gaussian distribution $\mathcal{N}(0,1)$. Then, the $W$ matrix is scaled according to the defined spectral radius, a hyperparameter defining the highest eigen value of $W$.

The final layer is a linear regression with ridge penalization where the
explanatory features are the reservoir state and the variable to be
explain is the outcome to predict such that:

$$W_{out} = YX^T ( XX^T + \lambda  I)^{ -1 }$$

Where x(t) and y(t) are accumulated in X and Y respectively such that:

$$X = \begin{bmatrix} x(1) \\ x(2) \\ ... \\ x(T) \end{bmatrix}
\text{ and } Y = \begin{bmatrix} y(1) \\ y(2) \\ ... \\ y(T) \end{bmatrix}$$

The parameter $\lambda$ is the ridge penalization which aims to prevent overfitting. Additionally, one can also connect the input layer to the output layer to the reservoir nodes. In that case, $X$ is the accumulation of both such that :

$$X = \begin{bmatrix} x(1), u(1) \\ x(2), u(2) \\ ... \\ x(T), u(T) \end{bmatrix}
\text{ and } Y = \begin{bmatrix} y(1) \\ y(2) \\ ... \\ y(T) \end{bmatrix}$$

Overall, there are four main hyperparameters to be chosen by the user: the leaking rate which defines the memory of the RC, the input scaling which define the relative importance of the features, the spectral radius which define the connections of the neurons inside the reservoir which in turn define the degree of non-linear combination of features and the ridge penalization which controls the degree of overfitting. The choice of hyperparameter often requires the user to evaluate the performance of different combinations of hyperparameters on a validation set before selecting the optimal combination to forecast on the test set.

# Basic package use

In this section, we will cover the basics of reservoirnet use including
installation, classification and regression. We will provide brief
overview of the package use and more in depth description is provided in
section 4 with the covid-19 forecast use case.

## Installation

reservoirnet is an R package api making the python module reservoirPy
easily callable from R. It is available on CRAN (see
<https://cran.r-project.org/package=reservoirnet>) and can be installed
using:

```{r eval=FALSE}
install.packages("reservoirnet")
```

Reservoir Computing (RC) is well suited to both regression and
classification tasks. We will introduce a simple example for both task.

## Regression {#basicregression}

### Covid-19 data

In this first use case, we will introduce the fundamental usage of the `reservoirnet` package. This demonstration will be conducted using the COVID-19 dataset that is included within the package. These data encompass hospitalization, positive RT-PCR results, and overall RT-PCR data sourced from Santé Publique France, which are publicly available on data.gouv.fr (for further details, refer to `help(dfCovid)`). Our primary objective is to predict the number of hospitalized patients 14 days into the future. To accomplish this, we will initially train our model on data preceding the date of January 1, 2022, and subsequently apply it to forecast values using the subsequent dataset.

We can proceed by loading useful packages (i.e ggplot2 [@wickham_ggplot2_2016], dplyr [@wickham_dplyr_2023] and patchwork [@pedersen_patchwork_2023]), data and define the task:

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(patchwork)
library(reservoirnet)

data("dfCovid")
dist_forecast = 14
traintest_date = as.Date("2022-01-01")
```

Due to the substantial fluctuations observed in both RT-PCR metrics, our initial step involves applying a moving average computation over the most recent 7-day periods for these features. Additionally, we augment the dataset by introducing an `outcome` column and an `outcomeDate` column, which will serve as valuable inputs for model training. Moreover, we calculate the `outcome_deriv` as the difference between the outcome and the number of hospitalized patients (`hosp`), representing the variation in hospitalization in relation to the current count of hospitalized individuals. The resulting smoothed data is visualized in Figure \@ref(fig:covidintro).

```{r}
dfOutcome <- dfCovid %>%
  # outcome at 14 days
  mutate(outcome = lead(x = hosp, n = dist_forecast),
         outcomeDate = date + dist_forecast,
         outcome_deriv = outcome - hosp) %>%
  # rolling average for tested and positive_pcr
  mutate_at(.vars = c("Positive", "Tested"),
            .funs = function(x) slider::slide_dbl(.x = x,
                                                  .before = 6,
                                                  .f = mean))
```

```{r covidintro, fig.cap="Hospitalizations, IPTCC and positive PCR of Bordeaux University Hospital.", echo = FALSE}
dfOutcome %>%
  tidyr::pivot_longer(cols = c("hosp", "Positive", "Tested")) %>%
  ggplot2::ggplot(mapping = aes(x = date, y = value)) +
  geom_line() +
  facet_grid(name ~ ., scales = "free_y") +
  theme_minimal() +
  labs(color = "") +
  theme(legend.position = "bottom")
```

### First reservoir

Setting a reservoir is done with the `createNode()` function. The important hyperparameters are the following :

-   Number of nodes (`units`) : it corresponds to the number of nodes inside the reservoir. Usually, the more the better, but more nodes increases the computation time.
-   Leaking rate (`lr`) : the leaking rate corresponds to the balance between the new inputs and the previous state. A leaking rate of 1 only consider information from new inputs.
-   Spectral radius (`sr`): the spectral radius is the maximum absolute eigenvalue of the reservoir connectivity matrix. A small spectral radius induces stable dynamics inside the reservoir, a high spectral radius induces a chaotic regime inside the reservoir.
-   Input scaling (`input_scaling`): the input scaling is a gain applied to the input features of the reservoir.
- Warmup (`warmup`) : it corresponds to the number of time step during which the data are propagating into the reservoir but not used to fit the output layer. This hyperparameter is set in the `reservoirR_fit()` function.

In addition, we can set the seed (`seed`). Because the reservoir connections are set at random, setting the seed is a good approach to ensure reproducibility.

For this part of the tutorial, we will set the hyperparameter at a given value. Hyperparameter optimization will be detailed at section [Study case](#studycase).

```{r}
reservoir <- reservoirnet::createNode(nodeType = "Reservoir",
                                      seed = 1,
                                      units = 500,
                                      lr = 0.7,
                                      sr = 1,
                                      input_scaling = 1)
```

Then we can feed the data to the reservoir and see the activation state of the reservoir $x(t)$. To do so, we first prepare the data and transform it to a matrix:

```{r}
## select explanatory and transform it to an array
X <- dfOutcome %>%
  filter(outcomeDate < traintest_date) %>%
  select(hosp, Positive, Tested) %>%
  as.matrix()
```

Then we run the `predict_seq` function. It takes as input a node (i.e a reservoir or a reservoir associated with an output layer) and the feature matrix.

```{r}
reservoir_state <- predict_seq(node = reservoir, X = X)
```

Now we can visualize node activation using the `plot` function presented at figure \@ref(fig:nodeactivationbad).

```{r nodeactivationbad, fig.cap="20 random nodes activation over time."}
plot(reservoir_state)
```

Numerous nodes within the system exhibit a consistent equilibrium state. The challenge arises when the output layer attempts to extract knowledge from these nodes, as they do not convey meaningful information. This issue can be attributed to the disparate scales of the features. To address this concern, a practical approach involves normalizing the features by dividing each of them by their respective maximum values, thereby scaling them within the range of $-1$ to $1$ by dividing by the maximum of the absolute value. Of note, here the features will be scaled between $0$ and $1$ because all features are positive.

```{r}
stand_max <- function(x) return(x/max(abs(x)))
# scaled features
Xstand <- dfOutcome %>%
  filter(date < traintest_date) %>%
  select(hosp, Positive, Tested) %>%
  mutate_all(.funs = stand_max) %>%
  as.matrix() %>%
  as.array()
```

We then feed them to the reservoir and plot the node activation again. Compared to \@ref(fig:nodeactivationbad), the obtained node activation \@ref(fig:nodeactivationgood) shows interesting trend outputs as no node seems saturated.

```{r nodeactivationgood, fig.cap="20 random node activation over time. Scaled features."}
# feed them to the reservoir
reservoir_state_stand <- predict_seq(node = reservoir,
                                     X = Xstand,
                                     reset = TRUE)
# plot the output
plot(reservoir_state_stand)
```

### Forecast

In order to train the reservoir, we should train the last layer which
linearly combines the neuron's output.

#### Set the ESN

Initially, we establish the output layer, incorporating a ridge penalty set at `1e3`. It's important to note that this hyperparameter can be subject to optimization, a topic that will be explored in the forthcoming [Study Case](#studycase) section. This parameter plays a pivotal role in fine-tuning the model's conformity to the data. When set excessively high, the risk of underfitting arises, whereas setting it too low can lead to overfitting. We connect the output layer to the reservoir making the model ready to be trained.

```{r}
readout <- reservoirnet::createNode(nodeType = "Ridge",
                                    ridge = 1e3)
model <- reservoirnet::link(reservoir, readout)
```

#### Set the data

First we separate the train set on which we will learn the ridge coefficients and the test set on which we will make the forecast. We define the train set to be all the data before 2022-01-01 and the test data to be all the data to have forecast both on train and test sets.

```{r}
# train set
dftrain <- dfOutcome %>% filter(outcomeDate <= traintest_date)
yTrain <- dftrain %>% select(outcome)
yTrain_variation <- dftrain %>% select(outcome_deriv)
xTrain <- dftrain %>% select(hosp, Positive, Tested)
# test set
xTest <- dfOutcome %>% select(hosp, Positive, Tested)
```

We standardize with the same formula as seen before. We learn the standardization on the training set and apply it on the test set. Then we convert the dataframe to matrix.

```{r results=FALSE}
# copy train and test sets
xTrainstand <- xTrain
xTeststand <- xTest
# standardise based on training set values
ls_fct_stand <- apply(xTrain,
                      MARGIN = 2,
                      FUN = function(x) function(feature) return(feature/(max(x))))
lapply(X = names(ls_fct_stand),
       FUN = function(x){
         xTrainstand[,x] <<- ls_fct_stand[[x]](feature = xTrain[,x])
         xTeststand[,x] <<- ls_fct_stand[[x]](feature = xTest[,x])
         return()
       })
# convert to array
lsdf <- lapply(list(yTrain = yTrain,
                    yTrain_variation = yTrain_variation,
                    xTrain = xTrainstand,
                    xTest = xTeststand),
               function(x) as.matrix(x))
```

#### Train the model and predict

We then feed the reservoir with the train set. To do so, we set a `warmup` of `30` days during which the data are propagating into the reservoir but not used to fit the output layer.

```{r}
### train the reservoir ridge output
fit <- reservoirnet::reservoirR_fit(node = model,
                                    X = lsdf$xTrain,
                                    Y = lsdf$yTrain,
                                    warmup = 30,
                                    reset = TRUE)
```

Now that the ridge layer is trained, we can forecast. We set the parameter `reset` to `TRUE` in order to clean the reservoir from the data used by the training set.

```{r}
vec_pred <- reservoirnet::predict_seq(node = fit$fit,
                                      X = lsdf$xTest,
                                      reset = TRUE)
```

```{r fig.cap="Forecast"}
dfOutcome %>%
  mutate(pred = vec_pred) %>%
  na.omit() %>%
  ggplot(mapping = aes(x = outcomeDate)) +
  geom_line(mapping = aes(y = outcome,
                          color = "observed")) +
  geom_line(mapping = aes(y = pred,
                          color = "forecast")) +
  geom_vline(mapping = aes(color = "train-test sets",
                           xintercept = traintest_date)) +
  scale_color_manual(values = c("#3772ff", "#080708", "#df2935")) +
  theme_minimal() +
  labs(color = "", x = "Date", y = "Hospitalizations")
```

We observe that the model forecast is not fully accurate, both on the test set and the train set. In that case, one option could be to reduce ridge penalization to fit more closely the data, the optimization of ridge hyperparameter will be discussed at section [Study case](#studycase). Another possibility is to ease the learning of the algorithm by forecasting the variation of the hospitalization instead of the number of hospitalized patients. For that step, we will learn on the `outcome_deriv` contained in `yTrain_variation` data which is defined outcome as `outcome_deriv = outcome - hosp`. As depicted at figure \@ref(fig:figForecastDerivvsRaw), this strategy improved the model forecast.

```{r figForecastDerivvsRaw, fig.cap="Covid-19 hospitalizations forecast. The model is either trained to forecast the number of hospitalizations (denoted Raw) or the variation of the hospitalizations compared to current level of hospitalisation (denoted Variation)"}
## Fit reservoir on outcome variation instead of raw outcome
fit2 <- reservoirnet::reservoirR_fit(node = model,
                                     X = lsdf$xTrain,
                                     Y = lsdf$yTrain_variation,
                                     warmup = 30,
                                     reset = TRUE)
## Get the forecast on the test set
vec_pred2_variation <- reservoirnet::predict_seq(node = fit2$fit,
                                                 X = lsdf$xTest,
                                                 reset = TRUE)
## Transform the outome variation forecast into hospitalisation forecast
vec_pred2 <- vec_pred2_variation + xTest$hosp
## Plot the results
dfOutcome %>%
  mutate(Raw = vec_pred,
         Variation = vec_pred2) %>%
  tidyr::pivot_longer(cols = c(Raw, Variation),
                      names_to = "Outcome_type",
                      values_to = "Forecast") %>%
  na.omit() %>%
  ggplot(mapping = aes(x = outcomeDate)) +
  geom_line(mapping = aes(y = outcome,
                          color = "observed")) +
  geom_line(mapping = aes(y = Forecast,
                          color = "Forecast")) +
  geom_vline(mapping = aes(color = "train-test sets",
                           xintercept = traintest_date)) +
  facet_wrap(Outcome_type ~ .,
             labeller = label_bquote(cols = "Outcome" : .(Outcome_type))) +
  scale_color_manual(values = c("#3772ff", "#080708", "#df2935")) +
  theme_minimal() +
  labs(color = "", x = "Date", y = "Hospitalizations")
```

## Classification

### The Japanese vowel dataset

This example is largely inspired from the [classification tutorial of reservoirpy](https://github.com/reservoirpy/reservoirpy/blob/master/tutorials/5-Classification-with-RC.ipynb). To illustrate the classification task, we will use the Japanese vowel dataset (@kudo_multidimensional_1999). The data can be loaded from `reservoirnet` as follow :

```{r}
japanese_vowels <- reservoirnet::generate_data(dataset = "japanese_vowels")[[1]]
X_train <- japanese_vowels$X_train
Y_train <- japanese_vowels$Y_train
X_test <- japanese_vowels$X_test
Y_test <- japanese_vowels$Y_test
```

The dataset comprises 640 vocalizations of the Japanese vowel \ae, contributed by nine distinct speakers. Each vocalization represents a time series spanning between 7 and 29 time steps, encoded as a 12-dimensional vector denoting the Linear Prediction Coefficients (LPC). A visual representation of six distinct utterances from the test set, originating from three different speakers, is depicted in Figure \@ref(fig:vowelpresentation).

```{r vowelpresentation, fig.cap="Vowel dataset, sample with 3 speakers and 2 utterance each.", echo = FALSE}
vec_sample <- c(1, 2, 41, 42, 71, 72)
dfplot_vowel <- lapply(vec_sample,
                 FUN = function(i){
                   speaker <- which(Y_test[[i]] == 1)
                   X_test[[i]] %>%
                     as.data.frame() %>%
                     tibble::rowid_to_column(var = "Time") %>%
                     tidyr::pivot_longer(cols = -Time,
                                         names_to = "component",
                                         values_to = "LPC") %>%
                     mutate(speaker = speaker, .before = 1,
                            uterrance = i) %>%
                     return()
                 }) %>%
  bind_rows()

ggplot(dfplot_vowel, mapping = aes(x = Time, y = LPC, color = component)) +
  geom_line() +
  facet_wrap(uterrance ~ speaker,
             labeller = label_bquote(cols = "speaker" : .(speaker)),
             ncol = 2) +
  theme_minimal() +
  theme(legend.position = "none")

```

The primary objective involves the attribution of each utterance to its respective speaker, this is denoted as classification or sequence-to-vector encoding. The secondary objective involves the attribution of each time step of each utterance to its speaker, this is denoted as transduction or sequence-to-sequence encoding.

### Classification (sequence-to-vector model)

The first approach is the sequence-to-vector encoding. For this task we aim to predict the speaker of the whole utterance (i.e the label is assigned to the whole sequence). We first start by creating the reservoir and the output layer.

```{r}
reservoir <- reservoirnet::createNode("Reservoir", units = 500,
                                      lr=0.1, sr=0.9,
                                      seed = 1)
readout <- reservoirnet::createNode("Ridge",ridge=1e-6)
```

To perform this task, we need to modify the training and testing processes. Leveraging the inherent inertia of the reservoir, information from preceding time steps is preserved, effectively endowing the RC with a form of memory. Consequently, the final state vector encapsulates insights gathered from all antecedent states. In the context of the sequence-to-vector encoding task, only this ultimate state is employed. This process is executed as follows:

*XH : Markovian or “echo state” property ? c’est pour dire que les états finissent par être oubliés au bout d’un certain temps, donc c’est pas ce que tu veux dire là je pense*

*TF : j'ai remplacé par inertia qui parait effectivement plus approprié. Mon idée c'est que l'information s'accumule dans le réservoir au fur et à mesure qu'il "lit" la séquence de données.*

```{r}
states_train = list()
k <- 1
for (x in X_train) {
  states <- reservoirnet::predict_seq(node = reservoir, X = x,
                                      reset=TRUE)
  states_train[[k]] <- t(as.matrix(states[nrow(states),]))
  k <- k+1
}
```

Then we can train the readout based on this last state vector. In that case, `Y_train` contains a single label for each utterance.

```{r}
res <- reservoirnet::reservoirR_fit(readout,X = states_train, Y = Y_train)
```

The prediction is also modified using only the final state :

```{r}
Y_pred <- list()
k <- 1
for (x in X_test) {
  states <- reservoirnet::predict_seq(node = reservoir, X = x,
                                      reset=TRUE)
  y <- reservoirnet::predict_seq(node = readout,
                                 X = as.array(states[nrow(states),]))
  Y_pred[[k]] <- y
  k <- k+1
}
```

Figure \@ref(fig:seqtovec) shows the prediction for the 6 utterances depicted at figure \@ref(fig:vowelpresentation) where the model correctly identifies the speaker.

```{r seqtovec, fig.cap="Prediction in a sequence-to-sequence approach 6 samples with 3 speakers and 2 utterance each. The speaker to predict is depicted in blue. For each of the 6 utterance, the model correctly identifies the speaker."}
dfplotseqtovec <- lapply(vec_sample,
                 FUN = function(i){
                   speaker <- which(Y_test[[i]][1,] == 1)
                   Y_pred[[i]] %>%
                     as.data.frame() %>%
                     tidyr::pivot_longer(cols = everything(),
                                         names_to = "pred_speaker",
                                         values_to = "prediction") %>%
                     mutate(pred_speaker = gsub(x = pred_speaker,
                                                pattern = "V", "")) %>%
                     mutate(speaker = speaker, .before = 1,
                            uterrance = i,
                            target = speaker == pred_speaker) %>%
                     return()
                 }) %>%
  bind_rows()

ggplot(dfplotseqtovec,
       mapping = aes(x = pred_speaker,
                     y = prediction,
                     fill = target)) +
  geom_bar(stat = "identity") +
  facet_wrap(uterrance ~ speaker,
             labeller = label_bquote(cols = "speaker" : .(speaker)),
             ncol = 2) +
  scale_fill_manual(values = c("#BDBDBD", "#A3CEF1")) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(y = 'Score',
       x = "Speaker")
```

Then, we can also compute the overall accuracy :

```{r}
accuracy <- function(pred, truth) mean(pred == truth)

Y_pred_class <- sapply(Y_pred,
                       FUN = function(x) apply(as.matrix(x),1,which.max))
Y_test_class <- sapply(Y_test,
                       FUN = function(x) apply(as.matrix(x),1,which.max))

score <- accuracy(pred = Y_test_class, truth = Y_pred_class)

print(paste0("Accuracy: ", round(score * 100,3) ,"%"))
```

### Transduction (sequence-to-sequence model)

For this task, the goal is to predict the speaker for each time step of each utterance. The first step is to get the data where the label is repeated for each time step. This is easily done with the `repeat_targets` argument as follow :

```{r}
japanese_vowels <- reservoirnet::generate_data(
    dataset = "japanese_vowels",
    repeat_targets=TRUE)$japanese_vowels
X_train <- japanese_vowels$X_train
Y_train <- japanese_vowels$Y_train
X_test <- japanese_vowels$X_test
Y_test <- japanese_vowels$Y_test
```

Then we can train a simple Echo State Network to solve this task. For this example we will connect both the input layer and the reservoir layer to the readout layer which is performed by the `%>>%` operator :

```{r}
source <- createNode("Input")
readout <- createNode("Ridge",ridge=1e-6)
reservoir <- createNode("Reservoir",units = 500,lr=0.1, sr=0.9, seed = 1)

model <- list(source %>>% reservoir, source) %>>% readout
```

We can then fit the model and predict the labels for the test data. The `reset` parameter is set to `TRUE` to remove information from the reservoir from the training process.

```{r}
model_fit <- reservoirnet::reservoirR_fit(node = model,
                                          X = X_train,
                                          Y = Y_train,
                                          warmup = 2)

Y_pred <- reservoirnet::predict_seq(node = model_fit$fit,
                                    X = X_test,
                                    reset = TRUE)
```

From the `Y_pred` and `Y_test` we represent at figure \@ref(fig:figseqtoseq) the predictions for the same patients as in figure \@ref(fig:vowelpresentation).

```{r figseqtoseq, fig.cap="Prediction in a sequence-to-sequence approach 6 samples with 3 speakers and 2 utterance each. The higher the score of the speaker, the lighter the color.", fig.height=6}
dfplotseqtoseq <- lapply(vec_sample,
                 FUN = function(i){
                   speaker <- which(Y_test[[i]][1,] == 1)
                   Y_pred[[i]] %>%
                     as.data.frame() %>%
                     tibble::rowid_to_column(var = "Time") %>%
                     tidyr::pivot_longer(cols = -Time,
                                         names_to = "pred_speaker",
                                         values_to = "prediction") %>%
                     mutate(pred_speaker = gsub(x = pred_speaker,
                                                pattern = "V", ""),
                            speaker = speaker,
                            uterrance = i,
                            .before = 1) %>%
                     return()
                 }) %>%
  bind_rows()

ggplot(dfplotseqtoseq, mapping = aes(x = Time,
                                     y = pred_speaker,
                                     fill = prediction)) +
  geom_tile() +
  facet_wrap(uterrance ~ speaker,
             labeller = label_bquote(cols = "speaker" : .(speaker)),
             ncol = 2) +
  scale_fill_gradient2(low = "#8ECAE6", high = "#FB8500", mid = "#023047",
                       midpoint = 0) +
  theme_minimal() +
  labs(y = 'Predicted speaker',
       fill = "Prediction score")
```

For those 6 utterances, the model correctly identify the speaker for most of the time steps. We can then evaluate the overall accuracy of the model :

```{r}
Y_pred_class <- sapply(Y_pred, FUN = function(x) apply(as.matrix(x),
                                                       1,
                                                       which.max))
Y_test_class <- sapply(Y_test, FUN = function(x) apply(as.matrix(x),
                                                       1,
                                                       which.max))
score <- accuracy(array(unlist(Y_pred_class)), array(unlist(Y_test_class)))

print(paste0("Accuracy: ", round(score * 100,3) ,"%"))
```

# Study case: Covid-19 hospitalizations forecast {#studycase}

## Introduction

Since late 2020, millions of cases of SARS-CoV-2 infection have been documented across the globe [@carrat_evidence_2021, @world_health_organisation_who_2020, @noauthor_estimating_2022]. This ongoing pandemic has exerted significant strain on healthcare systems, resulting in a surge in hospitalizations. This surge, in turn, necessitated modifications to the healthcare infrastructure and gave rise to population-wide lockdown measures aimed at preventing the saturation of healthcare facilities [@kim_health_2020, @simoes_organisation_2021, @hubner_surgery_2020]. The capacity to predict the trajectory of the epidemic on a regional scale is of paramount importance for effective healthcare system management.

Numerous COVID-19 forecasting algorithms have been proposed using different methods (e.g ensemble, deep learning, compartmental), yet none has proven entirely satisfactory [@rahimi_review_2021, @cramer_evaluation_2022]. In France, short-term forecasts  with different methods have been evaluated with similar results [@pottier_forecast_2021, @paireau_ensemble_2022, @carvalho_analysis_2021, @mohimont_convolutional_2021]. In this context a machine learning algorithm based on linear regression with elastic-net penalization, leveraging both Electronic Health Records (EHRs) and public data, was implemented at Bordeaux University Hospital [@ferte_benefit_2022]. This model, which aimed at forecasting the number of hospitalized patients at 14 days, showed good performance but struggled to accurately anticipate dynamic shifts of the epidemic.

RC has been used in the context of covid-19 epidemic forecast [@ghosh_reservoir_2021, @kmet_bezier_2019, @liu_nanophotonic_2023, @ray_optimized_2021, @zhang_sentiment_2023]. Among them, @ghosh_reservoir_2021, @liu_nanophotonic_2023 and @ray_optimized_2021 used it to forecast epidemic, @zhang_sentiment_2023 performed sentiment analysis and @kmet_bezier_2019 used it to solve optimal control related to vaccine. The evaluation of RC for epidemic forecast showed promising results in all approaches, being competitive with Long-Short Term Memory (LSTM) and Feed-Forward Neural Network (FFNN) in @ray_optimized_2021. However, the test period was short for @ghosh_reservoir_2021 (21 and 14 days) and @ray_optimized_2021 (86 days) making it difficult to evaluate the behavior of the methods during epidemic dynamic shift. This was not the case for @liu_nanophotonic_2023 (6 months) but they implemented daily ahead forecast which would be difficult to use to manage a hospital. Finally, all three implementations used only one time series as input whereas it has been shown that using different data sources could improve forecast [@ferte_benefit_2022]. Therefore, it is still difficult to assess the usefulness of RC over a large period and using many time series as inputs.

RC can be viewed as an extension of penalized linear regression, where inputs undergo processing by a reservoir, introducing the capacity for memory and non-linear combinations. Given the effectiveness of penalized linear regression in COVID-19 forecasting, as highlighted in @ferte_benefit_2022, and the promising results exhibited by RC in epidemic forecasting, as demonstrated in studies such as @ghosh_reservoir_2021, @liu_nanophotonic_2023, and @ray_optimized_2021, we have opted to employ RC for the prediction of hospitalizations at 14 days at the University Hospital of Bordeaux.

The primary aim of this study is to assess the performance of RC in this forecasting task. Secondary objectives include (i) comparing the performance of RC with that of elastic-net penalized regression (identified as the optimal model in @ferte_benefit_2022) and (ii) evaluating variations in RC performance based on different architectural choices, such as the connection between the input layer and the output layer, and the use of one input scaling per feature versus a common input scaling.

## Methods

### Data

The study utilized aggregated data spanning from May 16, 2020, to January 17, 2022, regarding the COVID-19 epidemic in France, drawing from various sources to enhance forecasting accuracy. These sources encompassed epidemiological statistics from Santé Publique France, weather data from the National Oceanic and Atmospheric Administration (NOAA), both providing department-level data [@etalab_les_2020, @smith_integrated_2011] and Electronic Health Record (EHR) data from the Bordeaux Hospital providing hospital-level data. All data were daily updated. Santé Publique France data included information on hospitalizations, RT-PCR tests, positive RT-PCR results, variant prevalence, and vaccination data, categorized by age groups. NOAA data contributed temperature, wind speed, humidity, and dew point data, allowing for the computation of the COVID-19 Climate Transmissibility Predict Index [@roumagnac_etude_2021]. EHRs data included hospitalizations, ICU admissions, ambulance service records, and emergency unit notes, with relevant COVID-19-related concepts extracted from the notes. Data are discussed more in depth in @ferte_benefit_2022.

First derivative over the last $7$ days were computed to enrich model information. To take into account measurement error and daily noise variation, data were smoothed using a local polynomial regression with a span of 21 days. As previously described, input features were scaled between -1 and 1 by dividing the observed value by the maximum of the absolute value of the given input feature.

All data are publicly available. Weather data can be obtained from @smith_integrated_2011 using R package `worldmet` [@carslaw_worldmet_2023]. [Vaccine data](https://www.data.gouv.fr/fr/datasets/donnees-relatives-aux-personnes-vaccinees-contre-la-covid-19-1/) can be downloaded from @etalab_les_2020. EHRs data can be downloaded on dryad [@ferte_benefit_2023]. For privacy issues, publicly available EHRs data below 10 patients were obfuscated to 0. For convenience, all data were downloaded, merged and provided as replication material.

### Evaluation framework

The task was to forecast 14 days ahead the number of hospitalized patients. As seen at section [Regression](#basicregression), we will train the model to predict the variation of hospitalization, denoted as $hosp$, defined as $outcome_{t+14} = hosp_{t+14} - hosp_{t}$. Metrics computation and visualizations will be performed on the predicted number of hospitalizations denoted as $\widehat{hosp_{t+14}} = \widehat{outcome_{t+14}} + hosp_{t}$.

The dataset was separated into two periods. First period from May 16, 2020 to March 1, 2021 served to identify relevant hyperparameters. Remaining data was used to evaluate the model performance.

The performance of the model was evaluated according to several metrics:

- the mean absolute error : $MAE = mean(|\widehat{hosp_{t+14}}-hosp_{t+14}|)$.
- the median relative error : $MRE = median(|\frac{\widehat{hosp_{t+14}}-hosp_{t+14}}{hosp_{t+14}}|)$.
- the mean absolute error to baseline : $MAEB =  mean(|\widehat{hosp_{t+14}}-hosp_{t+14}| - |hosp_{t}-hosp_{t+14}|)$.
- the median relative error to baseline : $MREB = median(|\frac{\widehat{hosp_{t+14}}-hosp_{t+14}}{hosp_{t}-hosp_{t+14}}|)$)

Mean was chosen over median for $MRE$ and $MREB$ because those metrics tend to have extremely high values when the denominator is close to 0 (i.e when the number of hospitalized patients is close to 0 or the number of patients hospitalized at 14 days is close to the current number of hospitalized patients respectively). $MAEB$ and $MREB$ compare model performance to a baseline model which predicts the current number of hospitalized patients at 14 days. Those metrics help to determine the information added by the model and is a good baseline as covid-19 forecast model do not always outperform this basic forecast (@cramer_evaluation_2022).

Because the outcome is obfuscated below 10 hospitalizations for privacy reason, we set both the outcome and the forecast to 10 when the observed value was 0 or the forecasted value was below 10 when evaluating the model performance.

### Models

We compared RC to elastic-net penalized regression (denoted as Enet). Furthermore we evaluated RC based on several architectures. First we compared RC with a single input scaling common to all features and a RC with on specific input scaling per feature. Second we compared RC where the input layer is connected to the output layer in addition to the connection between reservoir and output layer. Therefore, five models were evaluated :

- Elastic-net penalized regression denoted *Enet*
- RC with a single input scaling and no connection between input and ouput layers denoted *Common IS R %>>% O*
- RC with a single input scaling and connection between input and ouput layers denoted *Common IS I+R %>>% O*
- RC with multiple input scaling and no connection between input and ouput layers denoted *Multiple IS R %>>% O*
- RC with multiple input scaling and connection between input and ouput layers denoted *Multiple IS I+R %>>% O*

Because of the randomness of the reservoir, we took the median forecast of 10 reservoir on the train set to evaluate the performance of a given hyperparameter set. On the test set we aggregated the forecast of 40 reservoir, each of them having one of the 40 best hyperparameter sets found on the train set. In addition, because covid-19 hospitalization is a non-stationary process, models were re-trained everyday using all previous days. To ease computation burden, only one day over two was used to find hyperparameters on the training set.

### Hyperparameter optimisation using random search

RC relies mainly on 4 hyperparameters including the leaking rate (i.e "memory" parameter), spectral radius (i.e "chaoticity" parameter), input scaling (i.e "feature gain" parameter) and ridge (i.e penalization parameter). Input scaling can be either, common to all features or specific to each feature which increases the number of hyperparameter by the number of features.

Following the notation from `glmnet` package [@friedman_regularization_2010], elastic-net penalized linear regression relies on two hyperparameters, lambda (i.e the penalization parameter) and alpha (i.e the compromise between lasso and ridge penalty)

Hyperparameter were selected in the training set (i.e before March 1, 2021) using a wrapper approach and a random search sampler using 2000 samples for each model. The sampling distribution were defined as follow :

- (RC) ridge and (Enet) lambda : log-uniform law defined between 1e-10 and 1e5
- (RC) input scaling and spectral radius : log-uniform law defined between 1e-5 and 1e5
- (RC) leaking rate : log-uniform law defined between 1e-3 and 1
- (Enet) alpha : uniform defined between 0 and 1

We provided large search space for all hyper-parameters. Search space was slightly reduced for leaking rate based on previous results and because a leaking rate of 1e-3 already imply that new inputs make the reservoir change really slowly which is not inline with the dynamic of covid-19 but would be appropriate for an application where the phenomena to forecast has a slow dynamic.

Finally, we provided an additional Enet model similar to the one in @ferte_benefit_2022 where alpha was set to $0.5$ and lambda was re-evaluated everyday in the test set based on previous data using the cross-validation procedure provided by `glmnet`.

## Results

```{r echo=FALSE}
ls_use_case <- readRDS(file = here::here("data/precomputed_results.rds"))
```

The goal of this task is to predict 14 days ahead the hospitalization. Figure \@ref(fig:ucPresentData) shows both the training set (i.e before 2021-03-01) and the test set where the blue curve correspond to the input features (first derivatives are not shown) and the orange curves correspond to the outcome the model is trained on (i.e the hospitalization variation) and the hospitalizations at 14 days on which the performance metrics are computed. The figures outline that the relation between the input features and the outcome evolve over time and that the time series is not stationary. For instance IPTCC seems correlated to the outcome except that it completely miss the summer 2021 increase.

```{r ucPresentData, echo = FALSE, fig.cap = "Covid-19 epidemic at BUH. Outcome of interest is presented in orange. Model is trained to forecast Outcome curve which corresonds to the difference between Hospitalisatiosn at 14 days and current hospitalisations. Other features are scaled (divide by the maximum of the feature) represented in darkblue.", fig.height=8}
ls_use_case$plot_present_data
```

### Hyperparameter selection

Figure \@ref(fig:ucHyperparam) shows the hyperparameter optimisation using random search for the different RC architectures. We observe that model with multiple input scaling achieved better performance on the train set compared to model with single input scaling which is expected as they can adapt more closely to the data thanks to specific input scaling for each feature.

As expected, we observe that the optimal leaking rate is above 1e-2 for all RC which is coherent with the short term dynamic of covid-19 epidemic. Trend for other hyperparameter are less clear even though best hyperparameter sets were close for RC with common input scaling and for RC with multiple input scaling.

```{r ucHyperparam, echo=FALSE, fig.cap="Hyperparameter evaluation on training set by random search. Hp sets with MAE above 30 were removed for clarity of visualisation.", fig.height=8}
ls_use_case$plot_reservoir_hp
```

Figure \@ref(fig:ucHyperparamMultipleIS) shows the hyperparameter search for RC with multiple input scaling and connected input layer. We observe that the random search tend to favor high importance given to derivative of positive RT-PCR (including the elderly) and the derivative of IPTCC. The rest of the feature do not show clear pattern.

```{r ucHyperparamMultipleIS, echo=FALSE, fig.cap="Hyperparameter evaluation on training set by random search of the model with multiple input scaling and no connection between input layer and output layer. Hp sets with MAE above 30 were removed for clarity of visualisation.", fig.height=8}
ls_use_case$plot_feature_input_scaling
```

### Forecast performance

```{r performance, echo = FALSE}
ls_use_case$table_perf_esn %>%
  knitr::kable(caption = "Model performance with several reservoir configuration. For each setting, 40 reservoir are computed and the forecast is the median of the 40 forecast. Results show the performance metrics : MAE = Mean Absolute Error, MRE = Median Relative Error, MAEB = Mean Absolute Error to Baseline, MREB = Median Relative Error to Baseline.",
               col.names = c("Model", "MAE", "MRE", "MAEB", "MREB"),
               booktabs = TRUE,
               linesep = c('', '\\addlinespace'), digits = 2)
```

Table \@ref(tab:performance) shows the performance on the test set. Best model according to all metrics was RC with common input scaling and connection between input and output layers. Having one input scaling per feature did not improve the model which might be due to low generalisability of the hyperparameter of the training set to the test set due to non-stationarity. Additionaly, connecting input layer to output layer improved the model forecast.

We observe that elastic-net did not perform well and stayed close to the baseline prediction. However, when the penalisation is updated each day, the peformance improve. This is emphasize the problem of the transferrability of the hyperparameters to which multiple input scaling models seem more sensitive as they have more hyperparameters.

Figure \@ref(fig:forecastwithorwithoutUpdate) shows the forecast of the different models. We note that models struggle to accurately forecast slope shifts. For instance, summer 2021 initial increase is partially predicted by all models but its decrease is not well predicted. Winter 2021 increase is anticipated by all models but they tend to overestimate it because of the rise of vaccine effect.

```{r forecastwithorwithoutUpdate, echo = FALSE, fig.cap = "Reservoir computing forecast depending on the setting with and without monthly update. Red line is the median forecast of 40 reservoirs. Grey lines are individual forecast of each of the 40 reservoir.", fig.height=8}
ls_use_case$plot_forecast
```

### Number of model to aggregate

Figures \@ref(fig:ucAggregationSpread) show the individual forecast for the 40 best set of hyperparameters of each RC architecture. Due to the internal random connection of the reservoir, we observe forecast stochasticity and relying on only one forecast is unreliable. We explored the number of model needed at figure \@ref(fig:ucAggregationNb) which shows that after 10 models, forecast is stable and even 5 models for the simpler model with common input scaling which rely on less hyperparamters.

```{r ucAggregationSpread, echo = FALSE, fig.cap = "Individual forecast the 40 best hyperparameter sets for the different RC configuration. Forecast value above 200 were set to 200 for clarity.", fig.height=6}
ls_use_case$plot_before_aggregation_forecast
```

```{r ucAggregationNb, echo = FALSE, fig.cap = "Mean absolute error depending on the number of aggregated reservoir.", fig.height=6}
ls_use_case$plot_aggregation
```

## Discussion

In this specific application, we have demonstrated that RC exhibits commendable performance in comparison to Elastic-net, which serves as the reference model. Furthermore, we highlight the inherent challenges in forecasting within this context, primarily stemming from the non-stationarity of the time series.

All computations in this study were conducted using the reservoirnet package, and the entire codebase is accessible on GitHub (**todo : open to the public**). The package demonstrates its efficacy in implementing various reservoir architectures, including connection between the input layer and the output layer, as well as the utilization of several input scaling, all within the context of a real-world use case.

Given the substantial number of hyperparameters involved, we acknowledge that random search may not be the most efficient optimization algorithm. We have retained this approach for the sake of simplicity in this tutorial paper; however, meta-heuristic approaches, particularly those utilizing evolutionary algorithms, may prove more efficient, especially when employing multiple input scaling [@bala_applications_2018].

This study represents a novel contribution to epidemic forecasting utilizing reservoir computing. Notably, previous literature predominantly focused on simpler problems characterized by fewer input features or shorter evaluation periods [@ghosh_reservoir_2021, @liu_nanophotonic_2023, @ray_optimized_2021]. Our findings underscore the potential of this approach for future epidemics, suggesting its potential to surpass more traditional epidemiological tools while maintaining a lightweight model structure compared to other recurrent neural networks (RNNs).

It is worth noting that all models, including the @ferte_benefit_2022 models, encounter challenges in accurately anticipating slope shifts, indicating the need for further investigation. Specifically, additional work is warranted to extend the application of RC to high-dimensional settings, building upon the insights gained from models based on a more extensive set of features.

# Discussion and conclusion

In this paper, we introduce the R package `reservoirnet`, which serves as a versatile tool for implementing reservoir computing based on `ReservoirPy`'s Python API. It offers flexibility in defining the reservoir architecture, including options for specifying connections between the input layer and the output layer, as well as variations in input scaling as demonstrated on a real-world use case.

We provided a comprehensive overview of the basic usage of the `reservoirnet` package through illustrative examples in regression and classification tasks. This introductory section serves as a foundation for R users, offering step-by-step guidance on constructing and training reservoir computing models using the package. By demonstrating the application of RC in both regression and classification scenarios, we aim to equip users with the essential knowledge and skills needed to harness the capabilities of reservoir computing for diverse tasks.

Drawing on the robust foundation of the `ReservoirPy` structure, a well-maintained library, this package inherits its reliability and longevity. We have focused on providing access to the fundamental features, building upon the strong base provided by `ReservoirPy.` Therefore, this initial version of `reservoirnet` must evolve in tandem with the growing understanding and adoption of reservoir computing within the R community.

*XH : Il faut rephrase mais je suis pas sur de ce qui tu veux dire.*

*TF : j'ai reformulé ce dernier paragraphe en espérant que ça soit mieux. Mon idée c'était d'expliquer qu'on a focalisé le package sur les fonctions principales mais que tout n'est pas dispo depuis R (par exemple on ne peut pas choisir comment sont générées les matrices de connection aléatoires du réservoir, on les a laissé aux paramètres par défaut).*

# Bibliography